My title\thanks{Code and data are available at: \url{https://github.com/abdullah-motasim/Forecasting-2024-US-Presidential-Election}.}

### 2024-11-04 paper/paper.tex
Added models

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Abdullah Motasim
  - Elizabeth Luong
thanks: "Code and data are available at: [https://github.com/abdullah-motasim/Forecasting-2024-US-Presidential-Election](https://github.com/abdullah-motasim/Forecasting-2024-US-Presidential-Election)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
table-of-contents: true
toc-depth: 2
bibliography: references.bib
editor: 
  markdown: 
    wrap: sentence
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(testthat)
library(arrow)
library(here)
library(rstanarm)
library(bayesplot)
library(gt)
library(kableExtra)
library(knitr)
library(webshot2)

analysis_data <- read_parquet(here("data/02-analysis_data/analysis_data.parquet"))
```

# Introduction

@@

This study seeks to forecast the 2024 U.S.
Presidential election outcome by developing a generalized linear model based on poll aggregation data.
Specifically, we evaluate predictors including demographics, polling trends, and historical voting patterns.
While the popular vote does not directly determine the presidential outcome due to the electoral college system, our model’s findings offer a clearer picture of voter preferences and potential trends in key states.
This paper outlines our modeling approach, data processing techniques, and results, with a final section discussing model limitations and implications for future election forecasting.

@@

Telegraphing paragraph: The remainder of this paper is structured as follows.
@sec-data....

# Data {#sec-data}

## Overview

The aggregated dataset containing the polling data for the 2024 U.S. election was sourced from FiveThrityEight ().
We use the statistical programming language R [@citeR] and packages the ...
to download, analyze, and model the data.

@@

-   **pollster_name** - The name of the polling organization that conducted the poll (e.g., YouGov, RMG Research).

-   **candidate** - The name of the candidate in the poll (e.g., Kamala Harris, Donald Trump).

-   **percentage** - The percentage of the vote or support that the candidate received in the poll (e.g., 51.0 for Kamala Harris, 48.0 for Donald Trump).

-   **party** - The political party of the candidate in the poll (e.g., DEM for Democrats, REP for Republicans).

-   **sample_size** - The total number of respondents participating in the poll (e.g., 2712).

-   **state** - The U.S. state where the poll was conducted or focused, if applicable.

-   **numeric_grade** - A numeric rating given to the pollster to indicate their quality or reliability (e.g., 3.0).

These variables were selected due to their direct relevance to assessing candidate support, polling reliability, and geographic voting trends.
Specifically, pollster_name and numeric_grade are crucial for tracking potential biases and variations across polling organizations, as different methodologies can impact results.
Candidate/Party and percentage provide direct indicators of support levels for each contender, which is essential for making accurate projections of election outcomes.
Also, it is worth noting that there were more Sample_size is key to assessing the reliability of each poll, as larger samples tend to yield more accurate representations of public opinion.
Lastly, state enables us to account for regional variations, crucial in a system like the U.S.
Electoral College, where state-level results play a determining role in the election.
Collectively, these variables offer a focused and comprehensive view of the polling landscape, supporting a more reliable aggregation and interpretation of the data.

## Measurement

@@

Note this is a general overview of the measurement method utilized by Emerson, a more thorough breakdown of measurement can be found in \[Appendix A\].

## Outcome variables

@@

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

# ggplot(penguins, aes(x = island, fill = species)) +
#   geom_bar(alpha = 0.8) +
#   scale_fill_manual(values = c("darkorange","purple","cyan4"),
#                     guide = "none") +
#   theme_minimal() +
#   facet_wrap(~species, ncol = 1) +
#   coord_flip()
```

@@

And also planes (@fig-planes).
(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

# analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))

# analysis_data |> 
#   ggplot(aes(x = width, y = length)) +
#   geom_point(alpha = 0.8) +
#   theme_minimal() +
#   labs(x = "Wing width (mm)",
#        y = "Wing length (mm)")
```

@@

## Predictor variables

@@

# Model

@@

Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

@@

$$\text{Support Percentage} = \beta_0 + \beta_1 \cdot \text{pollster} + \beta_2 \cdot \text{sample size} + \beta_3 \cdot \text{state}$$
Where:

  - $\beta_0$ is the intercept term

  - $\beta_1$ represents the effect of the pollster source.

  - $\beta_2$ captures the influence of the sample size on support percentage.

  - $\beta_3$ accounts for state-level variations.

## Explaination of Variables and Inclusions

@@

1. **Pollster (`pollster`)**: Captures potential systematic differences among organizations, ensuring that support estimates are not bias by any single pollster's methodology.

2. **Sample Size (`sample_size`)**: Accounts for the reliability of each poll, as larger samples tend to reduce margin of error.

3. **State (`state`)**: Controls for geographic and political variations, allowing the model to adjust for differences in support based on regional preferences.

The summary statistics below @SummaryStatisticsComparison outline key variables in our analysis, focusing on Kamala Harris and Donald Trump. These tables display the distribution of support percentages and sample sizes, with metrics including mean, median, standard deviation, minimum, and maximum.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: SummaryStatisticsComparison
#| tbl-cap: Directly compares support percentages and sample sizes for Harris and Trump, showing central tendencies and variability in polling data for each candidate.

# Filter data for Harris and Trump
harris_data <- analysis_data %>%
  filter(candidate == "Kamala Harris")

trump_data <- analysis_data %>%
  filter(candidate == "Donald Trump")

# Define a function to compute summary statistics
compute_summary_stats <- function(data) {
  data %>%
    summarise(
      Mean = mean(percentage, na.rm = TRUE),
      Median = median(percentage, na.rm = TRUE),
      SD = sd(percentage, na.rm = TRUE),
      Min = min(percentage, na.rm = TRUE),
      Max = max(percentage, na.rm = TRUE),
      Sample_Size_Mean = mean(sample_size, na.rm = TRUE),
      Sample_Size_SD = sd(sample_size, na.rm = TRUE)
    )
}

# Summary statistics for Harris
harris_summary <- compute_summary_stats(harris_data) %>%
  mutate(Candidate = "Kamala Harris")

# Summary statistics for Trump
trump_summary <- compute_summary_stats(trump_data) %>%
  mutate(Candidate = "Donald Trump")

# Combine the two tables and format values to 2 decimal places
summary_table <- bind_rows(harris_summary, trump_summary) %>%
  select(Candidate, everything()) %>%
  mutate(across(where(is.numeric), ~ round(.x, 2)))  # Round numeric columns to 2 decimal places

# Rename columns in the summary table
summary_table <- summary_table %>%
  rename(
    `Sample Size Mean` = Sample_Size_Mean,
    `Sample Size SD` = Sample_Size_SD
  )

# Create a visual table using kable
summary_table %>%
  kable(caption = "Summary Statistics for Key Variables Comparing Kamala Harris and Donald Trump") %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position = "center") %>%
  add_header_above(c(" " = 1, "Percentage Statistics" = 5, "Sample Size Statistics" = 2)) %>%
  column_spec(2:8, width = "2em")  # Adjust column width if necessary

```

## Model justification

### Bayesian Framework

The models utilize a Bayesian approach with normal priors on the coefficients, implemented through the `rstanarm` package of @rstanarm in R [@citeR]. The Bayesian framework offers two primary advantages for our analysis:

1. **Uncertainty Representation**: Posterior distributions for each coefficient provide a clear indication of uncertainty around the estimates, allowing for more informative predictions.

2. **Regularization**: Normal priors help stabilize estimates, especially for smaller sample sizes or polls with limited data points, reducing overfitting.

### Monte Carlo Simulation for Win Probability Estimation

@@

1. **Generate Posterior Predictions**: Posterior distributions from each candidate’s model produce a range of possible support percentages for Harris and Trump.

2. **Simulate Election Outcomes**: Each simulation iteration compares Harris' and Trump’s predicted support. Harris is marked as the winner if her support surpasses Trump’s, and vice versa.

3. **Estimate Win Probabilities**: By repeating this comparison over thousands of iterations, we estimate win probabilities for each candidate based on the frequency of simulated wins.

## Model Assumptions and Limitations

@@

Second, the model assumes that each pollster’s effect on support estimates is consistent across all states, disregarding potential variations in polling methodologies and accuracy by region. This simplification may overlook regional nuances, especially if certain pollsters perform differently in specific states, potentially reducing accuracy in cases of state-specific polling practices.

Finally, the model treats state-level effects as static, assuming that state-specific support remains stable over time. This limits the model’s ability to adapt to changing voter sentiment within states as the election nears. Significant state-level shifts that occur after the latest data collection may not be reflected in the predictions, leading to outdated estimates.

@@

# Results

@@

## Predicted Support Percentages

### Kamala Harris

The distribution of predicted support for Kamala Harris is displayed in Figure 4. The model estimates her average support at approximately 47.32%, with the majority of values clustering around this mean. This relatively tight distribution indicates consistent support for Harris across polls, with limited variability. The standard deviation of her predicted support suggests that her polling results remain stable, even in states with more diverse polling outcomes.

### Donald Trump

Figure 5 illustrates the predicted support for Donald Trump, centered around a mean of 45.56%. Compared to Harris, Trump’s predicted support distribution shows slightly greater variability, indicating a broader spread in his support across polls. The higher standard deviation in Trump’s support highlights increased variability in polling outcomes, suggesting a wider range of voter responses depending on the region and poll characteristics.

### Comparison of Predicted Support

The combined density plot in Figure 6 provides a direct comparison of Harris and Trump’s predicted support distributions. Harris’s distribution (in blue) peaks at a higher level than Trump’s (in red), indicating her slight edge in predicted support. Both distributions overlap significantly, particularly between 45% and 50%, underscoring the close nature of their polling outcomes. This overlap suggests that while Harris holds a predicted lead, the margins are slim, indicating a competitive scenario in many states.

## Probability of Winning

Using Monte Carlo simulation, we estimated the likelihood of each candidate securing a win based on their predicted support. According to the model, Harris has a 63.2% probability of winning, while Trump has a 36.8% probability. These probabilities were calculated by comparing predicted support across thousands of iterations, with each iteration counting a win for the candidate who garnered higher support. This probabilistic approach captures the uncertainty inherent in election predictions and emphasizes Harris’ modest advantage in a hypothetical head-to-head match-up.

## State-Specific Trends

### Regional Variation in Support

The model reveals notable differences in support by state, aligning with known regional preferences. For instance, Harris’ support is consistently higher in traditionally Democratic states like California and New York, where her predicted vote share surpasses 60%. Conversely, Trump shows strong predicted support in states like Texas and Florida, with anticipated leads of 5-10 percentage points, reflecting historical Republican preferences.

### Battleground States

@@

## Summary of Predicted Insights

Overall, the model’s results showcase the impact of state-level factors, polling characteristics, and candidate-specific trends on predicted outcomes. The Bayesian GLM, enhanced with Monte Carlo simulations, enables the estimation of win probabilities and predicted support with a quantifiable degree of uncertainty. Harris’ predicted advantage, while modest, is supported by stable polling across states, whereas Trump’s broader distribution points to more variable support, especially in key states. This analysis provides a foundation for understanding competitive dynamics and regional differences in the 2024 U.S. presidential race, with implications for candidate strategies in battleground states.

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

# first_model <-
#   readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

# modelsummary::modelsummary(
#   list(
#     "First model" = first_model
#   ),
#   statistic = "mad",
#   fmt = 2
# )
```

# Discussion - Predictive Modeling of 2024 Election Outcomes

## Key Findings

The analysis utilizes Bayesian Generalized Linear Models (GLMs) and Monte Carlo simulations to estimate predicted support percentages for Kamala Harris and Donald Trump. With a win probability of 63.2% for Harris and 36.8% for Trump, the model suggests a modest advantage for Harris, albeit within a competitive margin. The density plot (Fig. 6) highlights the distributions of predicted support for both candidates, showing a slight overlap, which indicates close competition and underscores the uncertainty inherent in forecasting elections based on polling data.

@@

## Influence of Sample Size and Pollster Effects

@@

## State-Level Variations in Support

@@

## Weaknesses

@@

### Data-Related Weaknesses

@@

### Model-Related Weaknesses

The model’s reliance on a Gaussian distribution to predict support percentages may not fully account for the non-normality often observed in polling data. Polling data can exhibit skewed distributions or heavy tails, particularly in states where support for a candidate is highly polarized. By assuming a Gaussian distribution, the model may miss these nuances, potentially leading to biased estimates in states where the distribution of support is not symmetric.

@@

### Impact on Prediction Accuracy

These data and model weaknesses underscore the challenges of election forecasting and highlight the potential for misestimations in certain regions. The variability in polling data quality and the assumption of static state effects may reduce the model’s predictive accuracy, particularly in battleground states where small shifts in voter sentiment could have a significant impact on the election outcome. Furthermore, the lack of temporal adaptation in the model may lead to outdated predictions as the election approaches, particularly if there are last-minute shifts in public opinion.

## Future Directions

@@

## Conclusion

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={My title},
  pdfauthor={Abdullah Motasim; Elizabeth Luong},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{My title\thanks{Code and data are available at:
\url{https://github.com/abdullah-motasim/Forecasting-2024-US-Presidential-Election}.}}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{My subtitle if needed}
\author{Abdullah Motasim \and Elizabeth Luong}
\date{November 4, 2024}

\begin{document}
\maketitle
\begin{abstract}
First sentence. Second sentence. Third sentence. Fourth sentence.
\end{abstract}

\renewcommand*\contentsname{Table of contents}
{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{2}
\tableofcontents
}

\section{Introduction}\label{introduction}

@@

This study seeks to forecast the 2024 U.S. Presidential election outcome
by developing a generalized linear model based on poll aggregation data.
Specifically, we evaluate predictors including demographics, polling
trends, and historical voting patterns. While the popular vote does not
directly determine the presidential outcome due to the electoral college
system, our model's findings offer a clearer picture of voter
preferences and potential trends in key states. This paper outlines our
modeling approach, data processing techniques, and results, with a final
section discussing model limitations and implications for future
election forecasting.

@@

Telegraphing paragraph: The remainder of this paper is structured as
follows. Section~\ref{sec-data}\ldots.

\section{Data}\label{sec-data}

\subsection{Overview}\label{overview}

The aggregated dataset containing the polling data for the 2024 U.S.
election was sourced from FiveThrityEight (). We use the statistical
programming language R (R Core Team 2023) and packages the \ldots{} to
download, analyze, and model the data.

@@

\begin{itemize}
\item
  \textbf{pollster\_name} - The name of the polling organization that
  conducted the poll (e.g., YouGov, RMG Research).
\item
  \textbf{candidate} - The name of the candidate in the poll (e.g.,
  Kamala Harris, Donald Trump).
\item
  \textbf{percentage} - The percentage of the vote or support that the
  candidate received in the poll (e.g., 51.0 for Kamala Harris, 48.0 for
  Donald Trump).
\item
  \textbf{party} - The political party of the candidate in the poll
  (e.g., DEM for Democrats, REP for Republicans).
\item
  \textbf{sample\_size} - The total number of respondents participating
  in the poll (e.g., 2712).
\item
  \textbf{state} - The U.S. state where the poll was conducted or
  focused, if applicable.
\item
  \textbf{numeric\_grade} - A numeric rating given to the pollster to
  indicate their quality or reliability (e.g., 3.0).
\end{itemize}

These variables were selected due to their direct relevance to assessing
candidate support, polling reliability, and geographic voting trends.
Specifically, pollster\_name and numeric\_grade are crucial for tracking
potential biases and variations across polling organizations, as
different methodologies can impact results. Candidate/Party and
percentage provide direct indicators of support levels for each
contender, which is essential for making accurate projections of
election outcomes. Also, it is worth noting that there were more
Sample\_size is key to assessing the reliability of each poll, as larger
samples tend to yield more accurate representations of public opinion.
Lastly, state enables us to account for regional variations, crucial in
a system like the U.S. Electoral College, where state-level results play
a determining role in the election. Collectively, these variables offer
a focused and comprehensive view of the polling landscape, supporting a
more reliable aggregation and interpretation of the data.

\subsection{Measurement}\label{measurement}

@@

Note this is a general overview of the measurement method utilized by
Emerson, a more thorough breakdown of measurement can be found in
{[}Appendix A{]}.

\subsection{Outcome variables}\label{outcome-variables}

@@

Some of our data is of penguins (\textbf{?@fig-bills}), from Horst,
Hill, and Gorman (2020).

@@

And also planes (\textbf{?@fig-planes}). (You can change the height and
width, but don't worry about doing that until you have finished every
other aspect of the paper - Quarto will try to make it look nice and the
defaults usually work well once you have enough text.)

@@

\subsection{Predictor variables}\label{predictor-variables}

@@

\section{Model}\label{model}

@@

Background details and diagnostics are included in
Appendix~\ref{sec-model-details}.

\subsection{Model set-up}\label{model-set-up}

@@

\[\text{Support Percentage} = \beta_0 + \beta_1 \cdot \text{pollster} + \beta_2 \cdot \text{sample size} + \beta_3 \cdot \text{state}\]
Where:

\begin{itemize}
\item
  \(\beta_0\) is the intercept term
\item
  \(\beta_1\) represents the effect of the pollster source.
\item
  \(\beta_2\) captures the influence of the sample size on support
  percentage.
\item
  \(\beta_3\) accounts for state-level variations.
\end{itemize}

\subsection{Explaination of Variables and
Inclusions}\label{explaination-of-variables-and-inclusions}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Pollster (\texttt{pollster})}: Captures potential systematic
  differences among organizations, ensuring that support estimates are
  not bias by any single pollster's methodology.
\item
  \textbf{Sample Size (\texttt{sample\_size})}: Accounts for the
  reliability of each poll, as larger samples tend to reduce margin of
  error.
\item
  \textbf{State (\texttt{state})}: Controls for geographic and political
  variations, allowing the model to adjust for differences in support
  based on regional preferences.
\end{enumerate}

The summary statistics below (\textbf{SummaryStatisticsComparison?})
outline key variables in our analysis, focusing on Kamala Harris and
Donald Trump. These tables display the distribution of support
percentages and sample sizes, with metrics including mean, median,
standard deviation, minimum, and maximum.

\begin{longtable}[t]{l>{\raggedleft\arraybackslash}p{2em}>{\raggedleft\arraybackslash}p{2em}>{\raggedleft\arraybackslash}p{2em}>{\raggedleft\arraybackslash}p{2em}>{\raggedleft\arraybackslash}p{2em}>{\raggedleft\arraybackslash}p{2em}>{\raggedleft\arraybackslash}p{2em}}
\caption{Directly compares support percentages and sample sizes for Harris and
Trump, showing central tendencies and variability in polling data for
each candidate.}\\
\toprule
\multicolumn{1}{c}{ } & \multicolumn{5}{c}{Percentage Statistics} & \multicolumn{2}{c}{Sample Size Statistics} \\
\cmidrule(l{3pt}r{3pt}){2-6} \cmidrule(l{3pt}r{3pt}){7-8}
Candidate & Mean & Median & SD & Min & Max & Sample Size Mean & Sample Size SD\\
\midrule
Kamala Harris & 47.32 & 47.6 & 4.46 & 25 & 70 & 910.86 & 544.67\\
Donald Trump & 45.56 & 46.0 & 5.37 & 21 & 70 & 879.21 & 524.25\\
\bottomrule
\end{longtable}

\subsection{Model justification}\label{model-justification}

\subsubsection{Bayesian Framework}\label{bayesian-framework}

The models utilize a Bayesian approach with normal priors on the
coefficients, implemented through the \texttt{rstanarm} package of
Goodrich et al. (2022) in R (R Core Team 2023). The Bayesian framework
offers two primary advantages for our analysis:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Uncertainty Representation}: Posterior distributions for each
  coefficient provide a clear indication of uncertainty around the
  estimates, allowing for more informative predictions.
\item
  \textbf{Regularization}: Normal priors help stabilize estimates,
  especially for smaller sample sizes or polls with limited data points,
  reducing overfitting.
\end{enumerate}

\subsubsection{Monte Carlo Simulation for Win Probability
Estimation}\label{monte-carlo-simulation-for-win-probability-estimation}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Generate Posterior Predictions}: Posterior distributions from
  each candidate's model produce a range of possible support percentages
  for Harris and Trump.
\item
  \textbf{Simulate Election Outcomes}: Each simulation iteration
  compares Harris' and Trump's predicted support. Harris is marked as
  the winner if her support surpasses Trump's, and vice versa.
\item
  \textbf{Estimate Win Probabilities}: By repeating this comparison over
  thousands of iterations, we estimate win probabilities for each
  candidate based on the frequency of simulated wins.
\end{enumerate}

\subsection{Model Assumptions and
Limitations}\label{model-assumptions-and-limitations}

@@

Second, the model assumes that each pollster's effect on support
estimates is consistent across all states, disregarding potential
variations in polling methodologies and accuracy by region. This
simplification may overlook regional nuances, especially if certain
pollsters perform differently in specific states, potentially reducing
accuracy in cases of state-specific polling practices.

Finally, the model treats state-level effects as static, assuming that
state-specific support remains stable over time. This limits the model's
ability to adapt to changing voter sentiment within states as the
election nears. Significant state-level shifts that occur after the
latest data collection may not be reflected in the predictions, leading
to outdated estimates.

@@

\section{Results}\label{results}

@@

\subsection{Predicted Support
Percentages}\label{predicted-support-percentages}

\subsubsection{Kamala Harris}\label{kamala-harris}

The distribution of predicted support for Kamala Harris is displayed in
Figure 4. The model estimates her average support at approximately
47.32\%, with the majority of values clustering around this mean. This
relatively tight distribution indicates consistent support for Harris
across polls, with limited variability. The standard deviation of her
predicted support suggests that her polling results remain stable, even
in states with more diverse polling outcomes.

\subsubsection{Donald Trump}\label{donald-trump}

Figure 5 illustrates the predicted support for Donald Trump, centered
around a mean of 45.56\%. Compared to Harris, Trump's predicted support
distribution shows slightly greater variability, indicating a broader
spread in his support across polls. The higher standard deviation in
Trump's support highlights increased variability in polling outcomes,
suggesting a wider range of voter responses depending on the region and
poll characteristics.

\subsubsection{Comparison of Predicted
Support}\label{comparison-of-predicted-support}

The combined density plot in Figure 6 provides a direct comparison of
Harris and Trump's predicted support distributions. Harris's
distribution (in blue) peaks at a higher level than Trump's (in red),
indicating her slight edge in predicted support. Both distributions
overlap significantly, particularly between 45\% and 50\%, underscoring
the close nature of their polling outcomes. This overlap suggests that
while Harris holds a predicted lead, the margins are slim, indicating a
competitive scenario in many states.

\subsection{Probability of Winning}\label{probability-of-winning}

Using Monte Carlo simulation, we estimated the likelihood of each
candidate securing a win based on their predicted support. According to
the model, Harris has a 63.2\% probability of winning, while Trump has a
36.8\% probability. These probabilities were calculated by comparing
predicted support across thousands of iterations, with each iteration
counting a win for the candidate who garnered higher support. This
probabilistic approach captures the uncertainty inherent in election
predictions and emphasizes Harris' modest advantage in a hypothetical
head-to-head match-up.

\subsection{State-Specific Trends}\label{state-specific-trends}

\subsubsection{Regional Variation in
Support}\label{regional-variation-in-support}

The model reveals notable differences in support by state, aligning with
known regional preferences. For instance, Harris' support is
consistently higher in traditionally Democratic states like California
and New York, where her predicted vote share surpasses 60\%. Conversely,
Trump shows strong predicted support in states like Texas and Florida,
with anticipated leads of 5-10 percentage points, reflecting historical
Republican preferences.

\subsubsection{Battleground States}\label{battleground-states}

@@

\subsection{Summary of Predicted
Insights}\label{summary-of-predicted-insights}

Overall, the model's results showcase the impact of state-level factors,
polling characteristics, and candidate-specific trends on predicted
outcomes. The Bayesian GLM, enhanced with Monte Carlo simulations,
enables the estimation of win probabilities and predicted support with a
quantifiable degree of uncertainty. Harris' predicted advantage, while
modest, is supported by stable polling across states, whereas Trump's
broader distribution points to more variable support, especially in key
states. This analysis provides a foundation for understanding
competitive dynamics and regional differences in the 2024 U.S.
presidential race, with implications for candidate strategies in
battleground states.

Our results are summarized in Table~\ref{tbl-modelresults}.

\section{Discussion - Predictive Modeling of 2024 Election
Outcomes}\label{discussion---predictive-modeling-of-2024-election-outcomes}

\subsection{Key Findings}\label{key-findings}

The analysis utilizes Bayesian Generalized Linear Models (GLMs) and
Monte Carlo simulations to estimate predicted support percentages for
Kamala Harris and Donald Trump. With a win probability of 63.2\% for
Harris and 36.8\% for Trump, the model suggests a modest advantage for
Harris, albeit within a competitive margin. The density plot (Fig. 6)
highlights the distributions of predicted support for both candidates,
showing a slight overlap, which indicates close competition and
underscores the uncertainty inherent in forecasting elections based on
polling data.

@@

\subsection{Influence of Sample Size and Pollster
Effects}\label{influence-of-sample-size-and-pollster-effects}

@@

\subsection{State-Level Variations in
Support}\label{state-level-variations-in-support}

@@

\subsection{Weaknesses}\label{weaknesses}

@@

\subsubsection{Data-Related Weaknesses}\label{data-related-weaknesses}

@@

\subsubsection{Model-Related Weaknesses}\label{model-related-weaknesses}

The model's reliance on a Gaussian distribution to predict support
percentages may not fully account for the non-normality often observed
in polling data. Polling data can exhibit skewed distributions or heavy
tails, particularly in states where support for a candidate is highly
polarized. By assuming a Gaussian distribution, the model may miss these
nuances, potentially leading to biased estimates in states where the
distribution of support is not symmetric.

@@

\subsubsection{Impact on Prediction
Accuracy}\label{impact-on-prediction-accuracy}

These data and model weaknesses underscore the challenges of election
forecasting and highlight the potential for misestimations in certain
regions. The variability in polling data quality and the assumption of
static state effects may reduce the model's predictive accuracy,
particularly in battleground states where small shifts in voter
sentiment could have a significant impact on the election outcome.
Furthermore, the lack of temporal adaptation in the model may lead to
outdated predictions as the election approaches, particularly if there
are last-minute shifts in public opinion.

\subsection{Future Directions}\label{future-directions}

@@

\subsection{Conclusion}\label{conclusion}

The analysis demonstrates that both pollster reliability and sample size
are critical factors in accurately predicting support percentages, while
state-specific effects capture essential regional variations. However,
the model's assumptions and limitations suggest avenues for further
refinement. Incorporating dynamic, turnout-adjusted, and multi-source
data into future models could yield more precise election forecasts. In
sum, while this model provides a solid foundation for understanding
potential outcomes in the 2024 election, continuous refinement and
adaptation to new data sources and methods will enhance the reliability
of election prediction models.
